---
title: "Lobster Data Exploration"
author: "Paloma Cartwright"
date: "16/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

## Reading in the data set 

```{r}
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>% 
  clean_names() %>% 
  uncount(count)
```

## Finding counts and mean sizes by site and year 

Create a summary table that finds the total counts, mean carapace lengths by site and year 

```{r}
lobster_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(total_counts = n(), 
            mean_carapace_length = mean(size_mm, na.rm = TRUE))

ggplot(data = lobster_summary, aes(x = year, y = total_counts)) +
  geom_point(aes(color = site)) + geom_line(aes(color = site)) +
  labs(title = "Total Lobsters Observed by Year for Each Site", 
       x = "Year", 
       y = "Total Lobsters", 
       color = "Site")

```

## Find the proportion of legal lobsters for each site for 2020


```{r}
legal_size <- 79.76

lobster_data_2020 <- lobsters %>%
  filter(year == 2020) %>%
  mutate(legal = case_when(size_mm < legal_size ~ "No",
                           size_mm >= legal_size ~ "Yes")) %>%
  group_by(site, legal) %>%
  summarize(number = n())

ggplot(data = lobster_data_2020, aes(x = site, y = number)) +
  geom_col(position = "fill") + 
  aes(fill = legal) + 
  labs(title = "Proportion of Legal Lobsters in California dive sites in 2020", 
       x = "Site", 
       y = "Proportion", 
       fill = "Is Legal?")
```
NAPL and IVEE are in MPAs so the lobsters have more chances to grow to maturity under protection that lobsters outside of the marine protected areas.

## Task 3: Random Lobster Wrangling

#### filter() practice 

Create and store a subset from IVEE CARP and NAPL 
```{r}
ex_1 <- lobsters %>%
  filter(site == c("IVEE", "CARP", "NAPL"))
```

Create a subset that only contains lobsters observed in August 
```{r}
ex_2 <- lobsters %>% 
  filter(month == 8)
```

```{r}
ex_3 <- lobsters %>% 
  filter(site == "AQUE" | size_mm > 70)
```

```{r}
ex_4 <- lobsters %>% 
  filter(!(site == "NAPL"))
```

#### group_by() %>% summarize() practice

```{r}
ex_5 <- lobsters %>% 
  group_by(site) %>% 
  summarize(mean_lobster_carapace = mean(size_mm, na.rm = TRUE), 
            std_lobster_carapace = sd(size_mm, na.rm = TRUE))

```

```{r}
ex_6 <- lobsters %>% 
  group_by(site, month) %>% 
  summarize(max_length = max(size_mm, na.rm = TRUE))
```

#### mutate() practice

```{r}
ex_7 <- lobsters %>% 
  mutate(size_cm = size_mm/10) %>% 
  mutate(site = tolower(site)) %>% 
  mutate(area = as.character(area))
```

#### case_when() practice 

```{r}
ex_8 <- lobsters %>%
  mutate(size_bin = case_when(size_mm <= 70 ~ "small",
                              size_mm > 70 ~ "large")) %>% 
  mutate(designation = case_when(site %in% c("IVEE", "NAPL") ~ "MPA", 
                                 !(site %in% c("IVEE", "NAPL")) ~ "not MPA"))
```

 
 
